---
title: "Plot methods applied to `SensorArray` class"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r chunk_opt, echo = FALSE}
opts_chunk$set(fig.path = "figure/")
```


## Libraries

```{r inc}
library(chemosensors)
```

## Default plot method

```{r sa_default_plot, cache = TRUE}
sa <- SensorArray(nsensors = 17)
plot(sa) # the same as `plotPolar(sa)`
```

The default plot method for `SensorArray` object is `plotPolar`.
The concentration values come from `concMax` method.

```{r}
concMax(sa)
```

## Custom code to replicate polar plot

```{r polar}
set <- c("A", "B", "C")
sc <- Scenario(set)
conc <- getConc(sc)
conc
```

Note that the values in `conc` concentration matrix are the same as output of `concMax` method.

Now let's get the sensor array `sa` response.
The noise has to be suppressed, in order to replicate the results of `plotPolar`.

```{r, cache = TRUE}
sa0 <- sa
nsd(sa0) <- 0
sdata <- predict(sa0, conc)
sdata
```

Now we need to prepare a data.frame for plotting.

```{r}
df <- sdata.frame(sa, conc = conc, sdata = sdata, feature = "step")
df
```

Note that the same sensor array data might by obtained by an appropriate indexing of `sdata` matrix.

```{r}
ind <- seq(2, 6, by = 2)
ind
sdata[ind, ]
```

Do some data manipulation.

```{r}
snames <- grep("^S", names(df), value = TRUE)
df <- df[, c(snames, "lab")]
```

```{r melt}
mf <- melt(df, id.vars = "lab")
head(mf)
```

```{r lines}
p1 <- ggplot(mf, aes(variable, value, group = lab, color = lab)) + geom_line()
p1
```

```{r lines_polar}
p2 <- p1 + coord_polar()
p2
```

## Other plots 

### `plotResponse`

```{r}
if(packageVersion("chemosensors") < "0.7.10") {
  warning("plotResponse correctly works in chemosensors >= 0.7.10")
}
```

```{r}
idx(sa)
```

```{r plotResponse}
plotResponse(sa)
plotResponse(sa, sensor = 8)
plotResponse(sa, gas = c("A", "B"), sensor = 11)
```

### `plotBox`

```{r plotBox}
set <- rep(c("A 0.001", "A 0.01", "A 0.05", "A"), 5)
plotBox(sa, set = set)
plotBox(sa, set = set, scales = "free_y")
```


## R session info

```{r} 
sessionInfo()
```

